---
title: "Assignment 1: Linear Models"
author: "Due Monday, October 2 at 5pm"
format: html
editor: visual
---

# Instructions

You will submit an HTML document to Canvas as your final version.

You may work with **one** other person for this assignment.  Make sure both your names are on the HTML document, and you should **both** upload a copy of the assignment to Canvas.

Your document should show your code chunks/cells as well as any output.  Make sure that only relevant output is printed.  Do not, for example, print the entire dataset in your final knitted file.

Your document should also be clearly organized, so that it is easy for a reader to find your answers to each question.

There may be a small penalty for submissions that are difficult to read or navigate.


## Part One: Data Exploration

The dataset we will study for this assignment contains information about health insurance costs for individuals with no dependents (children) in the United States.  The information contained in the data is:

* Age of primary beneficiary

* Gender of primary beneficiary (only female, male recorded)

* Body mass index, providing an understanding of body, weights that are relatively high or low relative to height, objective index of body weight (kg / m ^ 2) using the ratio of height to weight, ideally 18.5 to 24.9

* Whether the beneficiary smokes

* The beneficiary's residential area in the US, northeast, southeast, southwest, northwest.

* Individual medical costs billed by health insurance


```{r, include = FALSE, eval = FALSE}
ins <- read_csv(here::here("Materials", "Assignments", "data", "insurance.csv"))

set.seed(13938)

splits <- ins %>% 
  filter(children == 0 ) %>%
  select(-children) %>%
  initial_split()

splits %>%
  training() %>%
  write_csv(here::here("Materials", "Assignments", "data", "insurance_costs_1.csv"))


splits %>%
  testing() %>%
  write_csv(here::here("Materials", "Assignments", "data", "insurance_costs_2.csv"))
```

You can find this data at: https://www.dropbox.com/s/bocjjyo1ehr5auz/insurance_costs_1.csv?dl=1


1. Read in the dataset, and display some summaries of the data. 

```{r, version = "answer_key", message = FALSE, warning = FALSE}
library(tidyverse)
library(tidymodels)
```


```{r, version = "answer_key", message = FALSE}
ins <- read_csv("https://www.dropbox.com/s/bocjjyo1ehr5auz/insurance_costs_1.csv?dl=1")

summary(ins)
glimpse(ins)

ins %>% count(smoker)
ins %>% count(region)
ins %>% count(sex)
```


2. Fix any concerns you have about the data.

```{r, version = "answer_key"}
ins <- ins %>%
  mutate(
    smoker = factor(smoker),
    sex = factor(sex),
    region = factor(region)
  )
```


3. Make up to three plots comparing the response variable (`charges`) to one of the predictor variables.  Briefly discuss each plot.

```{r, version = "answer_key"}
ins %>%
  ggplot(aes(x = age, y = charges)) +
  geom_point()

ins %>%
  ggplot(aes(x = region, y = charges, fill = smoker)) +
  geom_boxplot()

ins %>%
  ggplot(aes(x = region, y = charges, fill = sex)) +
  geom_boxplot()
```



## Part Two:  Simple Linear Models


1.  Construct a simple linear model to predict the insurance charges from the beneficiary's age.  Discuss the model fit, and interpret the coefficient estimates.

```{r, version = "answer_key"}
lin_reg_spec <- linear_reg() %>%
  set_mode("regression") %>%
  set_engine("lm")

ins_lm_fit <- lin_reg_spec %>%
  fit(charges ~ age, ins) 

ins_lm_fit$fit %>% summary()
```

%%%
version: answer_key


For a person that is 0 years old, the average charge is $3611.

For each additional year in a person’s age, the charges increase by $229 on average.

The p-value of the model is basically 0. That means this model is very significant, i.e., there is probably a relationship between the variables.

The r-squared value is 0.09938. This means 9.94% of the variance in charges are due to age.

%%%

2.  Make a model that also incorporates the variable `sex`.  Report your results.

```{r, version = "answer_key"}
ins_lm_fit_2 <- lin_reg_spec %>%
  fit(charges ~ age*sex, ins) 

ins_lm_fit_2$fit %>% summary()
```

%%%
version: answer_key

Insurace charges are on average $1805 higher for men than for women.

Among women, the insurance charges increase by $243 on average for every year of age.

Among men, the insurance charges increase by (243 - 30 = ) $213 on average for every year of age.

The p-value of the slope for age is near 0, but the p-value for the adjustments due to sex are high.  This suggests that there may not be a relationship between a person's designated sex and their insurance prices.

The r-squared value is 0.1006. This means 10.06% of the variance in charges is explained by sex and age.

%%%

3.  Now make a model that does not include `sex`, but does include `smoker`.  Report your results.

```{r, version = "answer_key"}
ins_lm_fit_3 <- lin_reg_spec %>%
  fit(charges ~ age*smoker, ins) 

ins_lm_fit_3$fit %>% summary()
```

%%%
version: answer_key

Insurance charges are on average $25,105 higher for smokers.

Among nonsmokers, the insurance charges increase by $258 on average for every year of age.

Among smokers, the insurance charges increase by (258 - 29 = ) $229 on average for every year of age.

The p-value of the slope for `age` and for `smokeryes` are near 0, suggesting that both these variables have an influenc on insurance prices.  The p-value of the coefficient for the interaction term is high, suggesting that the relationship between age and charges is not significantly different for smokers versus nonsomkers.

The r-squared value is 0.7606.  This means 76% of the variance in charges is explained by age and smoker status.  (Much higher than the previous models!)
%%%


4.  Which model (Q2 or Q3) do you think better fits the data?  Justify your answer by calculating the MSE for each model, and also by comparing R-squared values.

```{r, version = "answer_key"}
ins <- ins %>% 
  mutate(
    predictions_Q2 = predict(ins_lm_fit_2, ins)$.pred,
    predictions_Q3 = predict(ins_lm_fit_3, ins)$.pred
  )

ins %>%
  rmse(truth = charges,
      estimate = predictions_Q2)

ins %>%
  rmse(truth = charges,
      estimate = predictions_Q3)
```


%%%
version: answer_key

Based on the R-squared value being much higher, and the rmse being much lower, the model in Q3 is MUCH better.
%%%



## Part Three:  Multiple Linear Models


Now let's consider including multiple *quantitative* predictors.

1. Fit a model that uses `age` and `bmi` as predictors. (Do not include an interaction term between these two.) Report your results.  How does the MSE compare to the model in Part Two Q1?  How does the R-squared compare?

```{r, version = "answer_key"}
ins_lm_fit_4 <- lin_reg_spec %>%
  fit(charges ~ age + bmi, ins) 

ins_lm_fit_4$fit %>% summary()

ins %>% 
  mutate(
    predictions_Q4 = predict(ins_lm_fit_4, ins)$.pred
  ) %>%
  rmse(
    truth = charges,
    estimate = predictions_Q4
  )
```

%%%
version: answer_key

Insurance charges increase by $216 on average for every year of age.

Insurance charges increase by $283 on average for every point on the bmi scale.

Both these coefficients are statistically significant.

12% of the variance in charges can be explained by year and age.

The RMSE is higher than in Q3, and barely lower than in Q2.
%%%


2. Perhaps the relationships are not linear.  Fit a model that uses `age` and `age^2` as predictors.  How do the MSE and R-squared compare to the model in P2 Q1?

Hint:  You can do this by using the following formula:

```{r, eval = FALSE}
charges ~ poly(age, 2)
```


```{r, version = "answer_key"}
ins_poly_2 <- lin_reg_spec %>%
  fit(charges ~ poly(age, 2), ins) 

ins_poly_2$fit %>% summary()

ins %>% 
  mutate(
    predictions = predict(ins_poly_2, ins)$.pred
  ) %>%
  rmse(
    truth = charges,
    estimate = predictions
  )
```

%%%
version: answer_key

The R-squared and RMSE are almost identical to when we just used age.
%%%



3. Fit a polynomial model of degree 4.  How do the MSE and R-squared compare to the model in P2 Q1?

```{r, version = "answer_key"}
ins_poly_4 <- lin_reg_spec %>%
  fit(charges ~ poly(age, 4), ins) 

ins_poly_4$fit %>% summary()

ins %>% 
  mutate(
    predictions = predict(ins_poly_4, ins)$.pred
  ) %>%
  rmse(
    truth = charges,
    estimate = predictions
  )
```


%%%
version: answer_key

The R-squared and RMSE are slightly better than when we just used age.
%%%


4. Fit a polynomial model of degree 12.  How do the MSE and R-squared compare to the model in P2 Q1?

```{r, version = "answer_key"}
ins_poly_12 <- lin_reg_spec %>%
  fit(charges ~ poly(age, 12), ins) 

ins_poly_12$fit %>% summary()

ins %>% 
  mutate(
    predictions = predict(ins_poly_12, ins)$.pred
  ) %>%
  rmse(
    truth = charges,
    estimate = predictions
  )
```


%%%
version: answer_key

The R-squared and RMSE are a little better than when we just used age.
%%%


5. According to the MSE and R-squared, which is the best model?  Do you agree that this is indeed the "best" model?  Why or why not?

%%%
version: answer_key

Based on the R-squared and RMSE, the best model is the 12-degree polynomial.  However, this is a bad way to choose; models with more flexibility will *always* have a better R-squared and RMSE, but they are in danger of overfitting to the data.
%%%



6. Plot the predictions from your model in Q4 as a **line** plot on top of the scatterplot of your original data.  


```{r, version = "answer_key"}
ins %>% 
  mutate(
    predictions = predict(ins_poly_12, ins)$.pred
  ) %>%
  ggplot() +
  geom_point(aes(x = age, y = charges)) +
  geom_line(aes(x = age, y = predictions))

```



## Part Four:  New data

Great news!  We've managed to collect data about the insurance costs for a few more individuals.  You can find the new dataset here:  https://www.dropbox.com/s/sky86agc4s8c6qe/insurance_costs_2.csv?dl=1


Consider the following possible models:

* Only `age` as a predictor.

* `age` and `bmi` as a predictor.

* `age`, `bmi`, and `smoker` as predictors (no interaction terms)

* `age`, and `bmi`, with both quantitative variables having an interaction term with `smoker`  (i.e. the formula `~ (age + bmi):smoker`)

* `age`, `bmi`, and `smoker`as predictors, with both quantitative variables having an interaction term with `smoker` (i.e. the formula `~ (age + bmi)*smoker`)


For each model, *fit* the model on the **original data**.

```{r, version = "answer_key"}
ins_new <- read_csv("https://www.dropbox.com/s/sky86agc4s8c6qe/insurance_costs_2.csv?dl=1")

fit_1 <- lin_reg_spec %>%
  fit(charges ~ age, data = ins)

fit_2 <- lin_reg_spec %>%
  fit(charges ~ age + bmi, data = ins)

fit_3 <- lin_reg_spec %>%
  fit(charges ~ age + bmi + smoker, data = ins)

fit_4 <- lin_reg_spec %>%
  fit(charges ~ (age + bmi):smoker, data = ins)

fit_5 <- lin_reg_spec %>%
  fit(charges ~ (age + bmi)*smoker, data = ins)
```


Then, use the fitted model to *predict* on the **new data**.

```{r, version = "answer_key"}
ins_new <- ins_new %>%
  mutate(
    preds_1 = predict(fit_1, ins_new)$.pred,
    preds_2 = predict(fit_2, ins_new)$.pred,
    preds_3 = predict(fit_3, ins_new)$.pred,
    preds_4 = predict(fit_4, ins_new)$.pred,
    preds_5 = predict(fit_5, ins_new)$.pred,
  )
```


Report the MSE for each model's **new** predictions.  Based on this, which is the best model to use?

```{r, version = "answer_key"}
ins_new %>%
  rmse(
    truth = charges,
    estimate = preds_1
  )

ins_new %>%
  rmse(
    truth = charges,
    estimate = preds_2
  )

ins_new %>%
  rmse(
    truth = charges,
    estimate = preds_3
  )

ins_new %>%
  rmse(
    truth = charges,
    estimate = preds_4
  )

ins_new %>%
  rmse(
    truth = charges,
    estimate = preds_5
  )
```
%%%
version: original

%%%

%%%
version: answer_key

The last model had the best MSE, although the second to last was close.
%%%


Make a plot showing the residuals of your final chosen model.


```{r, version = "answer_key"}
ins_new %>% 
  mutate(
    resids = charges - preds_1 
  ) %>%
  ggplot() +
  geom_point(aes(x = age, y = resids))

```


## Challenge

Turn in a **separate HTML file** for your Challenge, if you choose to try it out.

Using any variables in this dataset, and any polynomial of those variables, find the model that best predicts on the **new data** after being fit on the **original data**.  

Make a plot showing the residuals of your final chosen model.

