---
title: "Midterm Review"
resource_files:
- appforthat.jpg
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightLines: yes
      highlightStyle: github
      countIncrementalSlides: false
      ratio: '16:9'

---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(tidymodels)
library(flair)
library(emo)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
library(xaringanExtra)
style_mono_light(
  base_color = "#26116c",
  text_bold_color = "#fd5e53",
  title_slide_text_color = "#fff8e7",
  background_color = "#fff8e7",
  header_font_google = google_font("Roboto"),
  text_font_google   = google_font("Roboto Condensed"),
  code_font_google   = google_font("Droid Mono")
)
```

```{css, echo = FALSE}
.red{ color: red; }
.blue{ color: blue; }
.huge {
  font-size: 200%;
}
.large {
  font-size: 150%;
}
.tiny {
  font-size: 50%;
}
```

---
class: center, middle

# Week 1: Linear Regression

---
# Hands-on skills


`r emo::ji("check")` How to **specify** a linear regression model.

`r emo::ji("check")` How to **fit** a linear regression model (possibly with polynomial terms)

`r emo::ji("check")` How to use the model to **predict** on a new dataset.

`r emo::ji("check")` How to calculate **metrics** like RMSE.

`r emo::ji("check")` How to calculate and plot **residuals**.

---
# Concepts

`r emo::ji("check")` How to **interpret** the **coefficient estimates**

`r emo::ji("check")` How to **select a model** based on theoretical measures and/or cross-validated metrics

`r emo::ji("check")` How to report and interpret the **R-squared** value.

`r emo::ji("check")` How to interpret the **MSE**

`r emo::ji("check")` How to check **residuals** for model fit.

---
# Quick Quiz

Interpret the coefficients and the R-squared value of the model below:

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(palmerpenguins)
library(tidyverse)
library(tidymodels)

lm(body_mass_g ~ (bill_length_mm + bill_depth_mm)*sex, data = penguins) %>% summary()
```


---
class: center, middle, inverse

# Week 2: K-Nearest-Neighbors Regression

---
# Hands-on skills


`r emo::ji("check")` How to **specify** a KNN model.

`r emo::ji("check")` How to **transform variables**

`r emo::ji("check")` How to **fit** a KNN model

`r emo::ji("check")` How to use the model to **predict** on a new dataset.

`r emo::ji("check")` How to calculate **metrics** like RMSE.

`r emo::ji("check")` How to calculate and plot **residuals**.

---
# Concepts

`r emo::ji("check")` Be able to **describe** the KNN process.

`r emo::ji("check")` Discuss **why** we might transform variables, and how that changes the model.

`r emo::ji("check")` Discuss the pros and cons of different choices of **K**.

`r emo::ji("check")` Discuss the pros and cons of KNN compared to **linear regression**

---
# Quick Quiz

In which of the following scenarios would *standardizing the predictors* be important for a KNN model? Defend your answer.

* Our predictors are all **categorical**.

* We only have one predictor

* Our predictors are *height in inches* and *weight in pounds*

* Our predictors are *height in inches* and *sex*

* Our predictors are *bill length in mm* and *bill depth in mm*

---
# Quick Quiz

Suppose we fit a KNN model with K = the number of observations in the dataset, minus one.

What are the pros and cons of this choice?

---
class: center, middle, inverse

# Week 2 and beyond: Cross-Validation

---
# Hands-on Skills

`r emo::ji("check")` Perform **v-fold cross-validation** on a model.

`r emo::ji("check")` Collect **metrics** from cross-validation, possibly including metrics that aren't the default in `collect_metrics`.

`r emo::ji("check")` Reference CV results for **model selection** and **tuning**

`r emo::ji("check")` Reference CV results for **model assessment**

---
# Concepts

`r emo::ji("check")` Understand **which situations** require cross-validation and which do not.

`r emo::ji("check")` Discuss **how** and **why** cross-validation addresses the problem of overfitting

`r emo::ji("check")` Be able to describe where **randomness** occurs in the cross-validation process.

`r emo::ji("check")` Figure out the sample sizes of the **test** and **training** sets from the choice of **v**

`r emo::ji("check")` Choose a **tuning parameter** value from cross-validated results.

--

`r emo::ji("x")` What is the "correct" choice of v

---
# Quick Quiz

In which of the following situations would you need to use cross-validation? Defend your answer.

* Deciding whether to use a KNN regression or linear regression

* Deciding between a simple linear regression with `bill_length_mm` as the only predictor, or a simple linear regression with `bill_depth_mm` as a predictor.

* Fitting a **final model**

* Reporting the metrics of your **final model**

* Tuning **K** in a KNN model.

* Calculating metrics of prediction on a **new dataset**.


---
class: center, middle, inverse

# Week 3: Classification

---
# Hands-on Skills

`r emo::ji("check")` Be able to **specify** and **fit** a KNN model for classification.

`r emo::ji("check")` Be able to **specify** and **fit** a Logistic Regression model for classification.

`r emo::ji("check")` Be able to **specify** and **fit** a LDA or QDA model for classification.

`r emo::ji("check")` Produce **predictions** from a classification model, including a **confusion matrix**

`r emo::ji("check")` Compute the following **metrics** for classification:  *accuracy*, *precision*, *recall*, *ROC-AUC*, *Gini Index*.

--

`r emo::ji("x")` Logistic regression with more than two categories.

`r emo::ji("x")` Deep math behind ROC-AUC and Gini Index

`r emo::ji("x")` F1 score and other metrics

---
# Concepts

`r emo::ji("check")` **Interpret** the coefficients of a Logistic Regression model

`r emo::ji("check")` **Interpret** the discriminant loadings of a LDA model.

`r emo::ji("check")` Discuss a **confusion matrix**.

`r emo::ji("check")` Explain the **idea** of the following metrics: *accuracy*, *precision*, *recall*, *ROC-AUC*, *Gini Index*.

`r emo::ji("check")` Make an argument for why one metric might be a better choice than another for **model assessment** in various scenarios.

--

`r emo::ji("x")` Interpret a QDA model.

---
# Quick Quiz

Suppose we are trying to predict if a customer is likely to buy a particular product.  In the following descriptions, what metric is being referenced?

* "The chances that my model classifies a non-buyer as a buyer is 20%"

* "My model is not very sensitive to our choice of decision boundary."

* "The nodes of my decision tree are very pure."

* "My model correctly identifies about 75% of buyers."

* "The predictions from my model are correct 60% of the time."

---
# Quick Quiz

Discuss the following confusion matrix.  How good is our  model at classifying penguin species?

```{r, echo = FALSE}
knn <- nearest_neighbor(neighbors = 5) %>%
  set_mode("classification") %>%
  set_engine("kknn")

penguins <- penguins %>%
  mutate(
    species = factor(species)
  ) %>%
  drop_na(species, bill_length_mm)

knn_fit <- knn %>% fit(species ~ bill_length_mm, data = penguins)

penguins %>%
  mutate(
    preds = predict(knn_fit, penguins)$.pred_class
  ) %>%
  count(species, preds)
```



---
class: center, middle, inverse

# Week 4: Decision Trees

---
# Hands-on Skills


`r emo::ji("check")` Be able to **specify** and **fit** a single decision tree, a bagged tree model, and a random forest model.

`r emo::ji("check")` Produce **predictions** and **metrics** from any of these models.

`r emo::ji("check")` Make a **visualization** of a single decision tree.

---
# Concepts

`r emo::ji("check")` **Interpret** the visualization and/or output of a single decision tree fit.

`r emo::ji("check")` **Interpret** the output of a bagged tree (don't need to explain the numbers, just the idea of "most important variables")

`r emo::ji("check")` **Interpret** the output of a random forest (don't need to explain the numbers, just the idea of "most important variables")

`r emo::ji("check")` Explain the reasons for choosing a **bagged tree** or **random forest** model.

--
`r emo::ji("x")` Examine a single tree from a bagged tree or random forest fit.

---
# Quick Quiz

Interpret the following decision tree:

```{r, echo = FALSE}
library(rpart.plot)

tree <- decision_tree() %>%
  set_mode("classification") %>%
  set_engine("rpart")

tree_fit <- tree %>% fit(species ~ ., data = penguins)

rpart.plot(tree_fit$fit)
```

---
# Quick Quiz

Why might we prefer a **random forest** in the previous analysis?

---
class: center, middle, inverse

# Week 5:  Bootstrapping

---
<!-- ## Hands-on -->

<!-- `r emo::ji("check")` Manually perform a reasonable number of **bootstrapped resamples** -->

<!-- ## Concepts -->

<!-- `r emo::ji("check")` **Explain** the advantages and reasons for bootstrapping. -->


<!-- `r emo::ji("check")` **Discuss** the results of a bootstrapped process. -->

<!-- --- -->
<!-- # Quick Quiz -->

<!-- What is a reason we might want to **bootstrap** a Logistic Regression? -->

---
class: center, middle, inverse

# Your turn
## As a group, use the **palmer penguins** dataset to create a question for the exam

(**library(palmerpenguins)**, then the dataset *penguins* will be auto-loaded)