---
title: "Project Guidelines"
author: "Stat 434"
output: rmdformats::readthedown
---

```{r, include = FALSE}
set.seed(492)
```

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, message = FALSE, warning = FALSE)
```

# Overview

Your final project is your chance to demonstrate your knowledge from Stat 434 in a real-world setting.  You have two options for the project:  a **prediction challenge** or a **method tutorial**.


#### Due: Sunday, June 6th at midnight

No extensions will be granted for projects, except in highly unusual extenuating circumstances.  Please start early!



# Option A: Kaggle Competition

If you chose to focus on **prediction** for your project, you will be entering a competition on the data hosting website Kaggle.com.

You can find a list of competitions here: https://www.kaggle.com/competitions

These competitions generally supply a *training set* for you to build your model, and a *validation set*  (without the response variable included).  You will upload your predictions for the *validation set* to Kaggle.


You may enter an *open* or *closed* Kaggle competition, it's completely your choice.  The open competitions have more limited options - but they also come with prizes if you place highly! You may not enter more than **three tries** into the competition.  Otherwise, I will consider this to be overfitting to the validation set.


For your project, you will turn in:

* Evidence that you successfully entered the competition; e.g., a screenshot of your team's results.

* A write-up or video walking through your modeling process.

 

# Option B: Tutorial

If you choose to focus on **concepts** for your project, you will create a tutorial explaining one of the concepts from class.

Your tutorial should cover one or two "lectures" worth of material; i.e., between 50% and 100% of a chapter in the textbook.

The target audience for your tutorial is anyone with experience equivalent to the prerequisites of Stat 434: statistics experience up to the level of basic regression.  You can also assume a basic familiarity with R  (or any programming language you use).

Important notes:

* Do not simply  mimic my material from lecture.  You should come up with your own approach to how to explain the modeling ideas.

* Do not simply walk through an example analysis.  You need to explain the ideas and statistical concepts behind the topic, not just show it in action.

* Do not create a *coding* tutorial.  There may be code included to demonstrate ideas, but you are not teaching R/`tidymodels`.  You are teaching the intuition and decision-making behind the methods.

* Do not reference anything done in class, or use datasets that were used in lecture/assignments.  You're starting fresh!


You will turn in to me a **write-up** or **video**


# Further Details

#### Style

For either project option, I will be grading very strictly for **professionalism**.  

That is to say, your submission should have a style, quality, and clarity that could reasonably be shared in a public place, such as a journal or a blog pots.

The following phrases should *never appear* in your work:

* "As we discussed in class, ..."

* "The data that we chose is...."

* "Like it says in the textbook, ..."

* "We tried [something] but we couldn't get it to work..."


Instead, you should write as though you are a professional data scientist sharing your work with the world.

This doesn't mean your project needs to be stuffy and formal; you can use casual language, and include funny images or references, as long as this does not detract from the clarity of the writeup.

A good model for a clear writeups and videos can be found in the blog on this site: https://juliasilge.com/.  (Although that blog is focused on teaching coding tools, whereas yours should be focused on your modeling procedure and/or the concepts behind your model.)

If you chose the video option for your modeling walkthrough and/or tutorial, it needs to be a true walkthrough of code and ideas.  Do not give me a video of you talking over PowerPoint slides!  Your grade will suffer.  


#### Format

There are no strict guidelines for format or length of your submissions.

Roughly, I would say 3-5 pages of writeup or 3-5 minutes of video is probably about right in most cases.  But this might vary - maybe you have a ton of visualizations and thus need more pages; maybe you are able to clearly show all the information in a more succinct way.  

Remember, usually less is more.  Longer submissions often mean that you haven't boiled down your work to the "takeaway messages".


#### Working in Groups

You may work in groups of **up to three** for this project.

However, you will be expected to submit three (or two) deliverables if you are a group of three (or two). 

These deliverables should be *themed* or *connected* in some way, not just two independent works.  For example, if you are working with one partner, you might:

* Enter a Kaggle competition, and use the data from that competition to create a tutorial for one of your models.

* Enter two Kaggle competitions that relate to the same domain (e.g., two biological data competitions)

* Write two tutorials, using the same example data for both and/or tying the concepts together.


Be aware that I will be grading group submissions as if all the work were truly collaborative, regardless of how you split up the work behind the scenes.  That is, if you work in a group of two, and your partner fails to complete their section, you both get a grade of at most 50% - the choice to work together is yours, and I will not be supervising group interactions.


**IMPORTANT:** If you opt to work in a group, make sure to add yourselves all to the same "Final Project Group" on Canvas **before** you upload your project.



# Challenge Points

You have the opportunity to earn **30 Challenge Points** by incorporating a topic not covered in the class.

(Optional lectures on some of these will be given in Week 10.)

This means you would need to teach that topic in your tutorial, or include that method in your Kaggle model.

Some possible topics include:

* Partial Least Squares Regression
* Splines
* Bayes Rule Classification
* Neural Networks
* Deep learning
* Survival analysis
* Time series analysis
* Multiple testing
* Naive Bayes
* GLMs (besides logistic)
* Bayesian additive regression trees
* Matrix completion
* Model-based Clustering
* Association Rules and Frequent Itemsets
* Manifold Learning
* Matrix Completion

You can also propose a topic for my approval, if you have an idea.

Note that Challenge Point bonuses are split among the group:  That is, if a group of two submits one project that incorporates a new topic and one project that does not, each group member receives 15 Challenge Points.



