---
title: "Cross-Validation"
author: "YOUR NAME HERE"
format: 
  html:
    theme: dark
    code-fold: true
    code-line-numbers: true
    code-tools: true
    self-contained: true
editor: visual
execute:
  message: false
---

## Setup

Declare your libraries:

```{r}
#| label: libraries-r
#| include: false
library(tidyverse)
library(tidymodels)
```

## Code from slides

```{r}
ins <- read_csv("https://www.dropbox.com/s/bocjjyo1ehr5auz/insurance_costs_1.csv?dl=1")
head(ins)
```

```{r}
lr_mod <- linear_reg() %>%
  set_mode("regression") %>%
  set_engine("lm")
```

```{r}
ins_cvs <- vfold_cv(ins, v = 10)

bmi_poly_1_cv <- lr_mod %>%
  fit_resamples(charges ~ bmi,
                resamples = ins_cvs)

bmi_poly_2_cv <- lr_mod %>%
  fit_resamples(charges ~ poly(bmi, 2),
                resamples = ins_cvs)
```

```{r}
bmi_poly_1_cv %>% collect_metrics()

bmi_poly_2_cv %>% collect_metrics()
```

## Your turn

Suppose you want to fit a regression that predicts `charges` from `age` and `bmi`.

Consider four models:

**Model 1**: `age` and `bmi` both have polynomial degree 1 (linear) 

**Model 2**: `age` has polynomial degree 1, `bmi` has polynomial degree 2 (quadratic and linear) 

**Model 3**: `age` has polynomial degree 2, `bmi` has polynomial degree 1 

**Model 4**: `age` and `bmi` both have polynomial degree 2


(No interactions term for any model.)

Which model is best according to a **cross-validation study** with 10 splits?
